---
title: "Presentation Ninja"
subtitle: "&#x2694;&#xFE0F; xaringan +<br/>&#x1F60E; xaringanthemer"  
author: 
  - "Yihui Xie"
  - "Garrick Aden-Buie"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```
## Motivation

- Getting more experience working with real data and practicing EDA techniques

- **Exploring patterns** within the data and **finding relationships** between variables


---
## WTA Match Data

Women's Tennis Association match data from 2018 to 2021

- Each observation is an individual match

- Player statistics are separated by winner and loser

- Focused on aces, first serves, and court surface

---

## Hypotheses

1. Hard courts will have more aces than other surfaces

2. Match minutes will be longer in the final and semifinal rounds compared to earlier rounds

3. Players will excel on one surface type rather than all three
